---
title: box office
author: ''
date: '2021-02-15'
slug: box-office
categories: []
tags: []
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r Import Data}
library(tidyverse)
library(rvest)
library(broom)
library(lubridate)
library(plotly)
library(DT)
library(RColorBrewer)
library(WDI)
library(ineq)

theme_set(theme_light())



indicators <- c("cpi_series" = "FP.CPI.TOTL")
wdi_data <- WDI(indicator = indicators, country = "US")


# base_url <- "https://www.the-numbers.com/movie/budgets/all/"
# movie_vec <- seq(1, 6001, 100)
# movie_df <- data.frame()
# 
# for (i in  movie_vec) {
#   
#   movie_url <- paste0(base_url, i)
#   movie_html <- read_html(movie_url)
#   temp_df <- html_table(movie_html, fill = T) %>% 
#     as.data.frame() %>% 
#     select(-1)
#   movie_df <- bind_rows(movie_df, temp_df)
#   
# }
# 
# openxlsx::write.xlsx(movie_df, "BoxOffice.xlsx")
movie_df <- readxl::read_excel("C:/Users/Alpha Go/Documents/R/Machine Learning/4. BoxOffice/BoxOffice.xlsx")
```

```{r Clean Data}
data <- movie_df %>% 
  janitor::clean_names() %>% 
  rename(release_date_temp = release_date) %>% 
  mutate(release_date = mdy(release_date_temp),
         release_year_temp = year(release_date),
         release_year = ifelse(is.na(release_year_temp), parse_number(release_date_temp), release_year_temp),
         release_month = as.integer(month(release_date)),
         release_decade = as.factor(release_year - release_year %% 10),
         production_budget = parse_number(production_budget),
         domestic_gross = parse_number(domestic_gross),
         worldwide_gross = parse_number(worldwide_gross)) %>% 
  filter(worldwide_gross != 0 & release_date_temp != "Unknown" & release_year < 2020) %>% 
  filter(!(production_budget > 1e4 & worldwide_gross < 1000)) %>% 
  mutate(release_decade = fct_lump(release_decade, 4, other_level = "before_1980")) %>% 
  inner_join(wdi_data, by = c("release_year" = "year")) %>% 
  mutate(production_budget_adj = production_budget * (100/cpi_series),
         domestic_gross_adj = domestic_gross * (100/cpi_series),
         worldwide_gross_adj = worldwide_gross * (100/cpi_series),
         production_budget_adj_log = log(production_budget_adj),
         domestic_gross_adj_log = log(domestic_gross_adj),
         worldwide_gross_adj_log = log(worldwide_gross_adj)) %>% 
  mutate(across(ends_with("adj"), ~ ./1e6))
```

# Introduction
In the last couple of years a lot of people have noted that the share of high budget movie production has increased and that movies with smaller and mid-sized budgets have a more difficult time to obtain financing. By leveraging on data compiled by the website [The Numbers](https://www.the-numbers.com/movie/budgets/all) I will try to provide some insights on this issue. In addition, the dataset can help us to tackle some other questions as well.

1. Profitability
    - How has profitability evolved over time and what are some of most and least profitable movies.
2. Domestic vs Worldwide Gross
    - Overall how important is the domestic US market for movies and has this changed over time.
3. Predictive ability of movie budgets for box office success
    - Do more expensive movies make more money at the box office and how strong is this relationship.
4. Inequality in box office success
    - Does potential inequality in production budgets also translate into more inequality at the box office.

# Overview {.tabset .tabset-pills}

Overall, the final dataset contains `r nrow(data)` movies with data on Production Budgets as well as Domestic and Worldwide Gross measured in Millions (Mio.) of Dollars ($). Since the movies are spread across 6 decades the values are adjusted for inflation. Given that the dataset does not contain information on where a movie was produced the US consumer price index (CPI) is used to adjust for inflation.

The graph below plots the number of movies by release year. Starting in the 2000s we have data on about 200 movies per year. However, this number drops significantly before the late 1990s. For example, we have about 30 movies per year for the 1980s and even less before that. This suggests that the data before the late 1990 is likely restricted to more well known and more successful movies. This will likely bias some of the comparisons and should be kept in mind when interpreting the results.

The other tab depicts the relative frequency of movie releases by month. While there is some heterogeneity between decades January appear to be the least popular month for releasing movies. This effects appears to be even more pronounced for the 80s. At the same time, the share of releases in June is conspicuously high. One potential explanation for this might go as follows: 1) The data for the 80s is biased towards more popular movies and these are more likely to be released in summer. Whether or not this is true has to be settled another time. Alternatively there is always the option to call on the the elderly to ascertain the validity of any given chain of reasoning and provide an authoritative answer.

## Number of Movies over Time

```{r Number of Movies}
data %>% 
count(release_year) %>% 
ggplot(aes(release_year, n)) +
geom_line(col = "royalblue") +
scale_x_continuous(breaks = seq(1960, 2020, 10)) +
labs(title = "Number of Movies over time",
     x = "Release Year",
     y = "Number")


```
